Real Time Fluid Simulation \\ \smaller using Smoothed-Particle Hydrodynamics and OpenGL

### 2014-12-14 paper/paper.tex
Added Overview section

\section{Smoothed Particle Hydrodynamics Overview}
%

%

>>>>

\section{Smoothed Particle Hydrodynamics Overview}
%
\hspace{6 pt} Smoothed-particle hydrodynamics (SPH) method works by dividing the fluid into a set of 
referred to as particles. These particles have a spatial distance, or smoothing length, over which 
their properties are "smoothed" by a kernel function, such as a Gaussian function
\begin{align}
(\frac{1}{\pi h^2})^\frac{3}{2}  \exp{(\frac{-r^2}{h^2})}
\end{align}
This means that the physical quantity of any particle can be obtained by summing the 
relevant properties of all the particles which lie within the range of the kernel. For 
example, using Monaghan's popular cubic spline kernel the temperature at position r 
depends on the temperatures of all the particles within a radial distance 2h of r \cite{wikiarticle}.
\\
\hspace*{6 pt} The contributions of each particle to a property are weighted according to their distance 
from the particle of interest, and their density
A property for any one particle is calculated by discerning how many particles lay within
smoothing length of a given particle and what their respective properties are, and how 
dense that smoothing length is around the particle of interest. Mathematically, this is 
governed by a kernel function, like the Gaussian function given in eq. 1. This Gaussian
function gives large numbers for particles whose smoothing lengths are close to the 
particle of interest, and thus effect the calculation of a given particles properties, 
but as particles get far away, and their smoothing length becomes extremely small, 
although never reaches 0. This is due to intrinsic properties of the 
Gaussian function itself, but it is easy to account for this problem by testing for 
extremely small values close to zero, and disregarding those particles that are far away 
from the particle of interest as a result.
\\
\hspace*{6 pt} The equation for any quantity A at any point r is given by the equation
\begin{align}
A(r) = \sum_{j} m_j \frac{a_j}{p_j} W (|r - r_j|,h)
\end{align}
where m is the mass of particle j, A is the value of the quantity A for particle j, r 
denotes the position, p is the particle's density and W is a kernel function, such as the 
Gaussian function used by Gingold and Monaghan shown in eq. 1. The density of 
the particle can be computed in a variety of ways, such as the example equation
\begin{align}
p_i = p(r_i) = \sum_{j} m_j \frac{p_j}{p_j} W (|r_i - r_i|,h) = \sum_{j} m_j W(r_i - r_j,h)
\end{align}
Similarly, the spatial derivative of a quantity can be computed as follows
\begin{align}
\triangledown A(r) = \sum_{j} m_j \frac{a}{p_j} \triangledown W(|r - r_j|, h)
\end{align}
Although the size of the smoothing length can be fixed in both space and time, this does 
not take advantage of the full power of SPH. By assigning each particle its own smoothing 
length and allowing it to vary with time, the resolution of a simulation can be made to 
automatically adapt itself depending on local conditions. For example, in a very dense 
region where many particles are close together the smoothing length can be made relatively 
short, yielding high spatial resolution. Conversely, in low-density regions where 
individual particles are far apart and the resolution is low, the smoothing length can be 
increased, optimising the computation for the regions of interest. Combined with an 
equation of state and an integrator, SPH can simulate hydrodynamic flows efficiently \cite{wikiarticle}.
\\
\hspace*{6 pt} Given the principles of smoothing length, calculating properties based on 
the area around a particle, and increasing or decreasing a property given the density of 
particles in an area, we applied some of these concepts in our implementation of SPH.
%
